{\bf \huge Wikiplia}:\\ The Free Programming Language\\ That Anyone Can Edit

### 2007-03-20 wikiplia/paper/wikiplia.tex
little comment
fix image
lparen, paper

\vspace{1em}
{\noindent \small {\bf Keywords}:
  Freedom, programming language, software license, wiki, XML, 
  weatherproof footwear and their fastening mechanisms,
  hyper-driven devices,
}

@@

\paragraph{Freedom \ftobe: Free to be You and Me.}
The development of a programming language should not be confined to
the bearded academic elite, gazing down upon the programmer fiefs from
their stratospheric ivory towers. Wikiplia is the free programming
language that {\em anyone} can edit: from bearded academic
elite\wcite{Tenure} to congressional staffers\wcite{Congressional
staffer edits to Wikipedia} to nameless Slashdot\wcite{Slashdot}
trolls. Wikiplia's WWW-based approach means that an Internet
connection and compatible WWW hypertext browser is all that's needed
to begin on the life-changing journey of programming language design.

@@

The hallmark of Free software is the GNU General Public License. It is
a hereditary license that requires that (a) source code be distributed
with the program and (b) modified versions of the program also be
licensed under the GPL. The intention is that anyone receiving the
software can exercise Freedom~\ftinker by understanding the source
code and modifying it to suit his needs. Clearly any source code will
not do: an obfuscated\wcite{Obfuscated code} version of the source
code cannot be easily understood or modified, even though it is
technically ``source code.'' The GPL therefore defines source code as
the ``preferred form'' for ``making modifications.''

@@

After this scenario, K and R believe they are working on the same
programming language---after all, it has the same source code---but
their minor bifurcation in development history means that they forever
differing meanings of the \verb+'\c'+ extension. This mistake is
likely to go unnoticed for some time, and until it is resolved, the
meaning of the \verb+'\c'+ extension is firmly enslaved in the
bipartite penitentiary of double entendre, yearning to be
free\ldots\z\wcite{Information wants to be free}

@@

Based on these scenarios we conclude that extant social measures such
as revision control conventions are not enough to guarantee freedom.
Even if we think these situations are implausible in the hands of
well-intentioned, well-mannered software engineers, we wish for our
software to remain free even when in the hands of nefarious and crafty
factions who would seek to fracture\wcite{Fork (software development)}
our free software community. We therefore need a technological and
legal solution that forces the entire development history to be
available. Keeping with Freedom~\fusa, we call this technology and its
license {\em Total Information Awareness} after the successful project
of the US Government with the same name.\z\wcite{Information Awareness
Office}

@@

The remainder of this paper proceeds as follows. We first discuss in
Section~\ref{sec:calculus} the design of the initial Wikiplia system,
which is used to bootstrap the rest of Wikiplia. We then discuss the
current state of Wikiplia as of revision \currentrevision\ in
Section~\ref{sec:current}. We discuss the freedom-preserving TIA
Public License in Section~\ref{sec:license}. We conclude with a
discussion of unrelated work and plans for the future~\ref{sec:conc}.

@@

\begin{figure*}[htp]
\[\begin{array}{cc}
% evaluation of lists
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf list}\ \vec{X} \m \tag{list}{\vec{X}}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf cons}\ X\ \tag{list}{\vec{X}} \m \tag{list}{X\ \vec{X}}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf lambda}\ \tag{symbol}{s}\ X \m \tag{closure}{\G\ s\ X}}{}} \\[1em]
\infer{\G \vdash \rate \tag{closure}{\G'\ s\ X}\ \vec{X} \m X'}{\G',s = \tag{list}{\vec{X}} \vdash \eval X \m X'} &
\infer{\G \vdash \rate \tag{prim}{\sf eval}\ X \m X'}{\G \vdash \eval X \m X'} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf xcase}\ \tag{list}{}\ X\ \vec{X} \m X'}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{list}{X_h \vec{X_t}}\  X_0\ \\
                  & \tag{list}{\tag{symbol}{s_h}\ \tag{symbol}{s_t}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_h = X_h, s_t = X_t \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{quote}{X}\  X_0\ X_1\ \\
                  & \tag{list}{\tag{symbol}{s_q}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_q = X \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{string}{s}\  X_0\ X_1\ X_2\ 
                    X\ \vec{X} \m X' \\
\end{array}
}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{int}{s}\  X_0\ X_1\ X_2\ X_3\ 
                    X\ \vec{X} \m X' \\
\end{array}
}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{symbol}{s}\  X_0\ X_1\ X_2\ X_3\ X_4\ \\
                  & \tag{list}{\tag{symbol}{s_s}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_s = \tag{string}{s} \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{t}{\vec{X_t}}\  X_0\ X_1\ X_2\ X_3\ X_4\ X_5\ X \vec{X} \m X' \\
\end{array}
}{t = {\tt prim} \,\,or\,\, {\tt closure} & \G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf quote}\ {X} \m \tag{list}{X}}{}} \\[1em]
%
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf head}\ \tag{string}{s} \m X
}{{\tt cvs\ checkout}\ s = X}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf read}\ \tag{string}{s} \tag{int}{i} \m X
}{{\tt cvs\ checkout\ -\!r}\ i\ s = X}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf insert}\ \tag{string}{s} X \m \tag{int}{i}
}{{\tt cvs\ commit}\ s\ X = i}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf history}\ \tag{string}{s} \m \tag{list}{\vec{X}}
}{{\tt cvs\ log}\ s = \vec{X}}} \\[1em]

@@

  \begin{tabular}{r@{\,\,=\,\,}l}
   \lb {\tt(} $E_1\ \ldots\ E_n$ {\tt)}\rb & $\tag{list}{ \lb E_1 \rb\ \ldots\ \lb E_n \rb}$ \\
   \lb {\tt"}$s${\tt"}\rb & $\tag{string}{s}$ \\
   \lb $n$\rb & $\tag{int}{n}$ \\
   \lb $sym$\rb & $\tag{symbol}{sym}$ \\
   \lb {\tt'}$E$\rb & $\tag{quote}{E}$ \\
  \end{tabular}

@@

XML documents are universally parseable.\z\wcite{Parsing} However,
they are difficult to write and read. Therefore, as usual\wcite{RELAX
NG} we create a new syntax by which humans can write and read
documents and which the computer automatically parses and converts to
the easily parsable XML syntax and back to the new syntax, reducing
complexity.\z\footnote{For efficiency, the Wikiplia implementation is
optimized to lazily perform these translations, so that the document
is never represented in XML form.} This compact syntax is based upon
parentheses rather than tags: The XML document $\tag{list}{X_1\ X_2}$
is instead written ${\tt(}X_1\ X_2{\tt)}$. Note that the closing
parenthesis is not {\em named} as in XML, which makes parsing
difficult because the computer must {\em guess} which parenthesis
belongs to which other parenthesis. We therefore call this compact
syntax XESP because it basically {\em reads the programmer's
mind}\,\wcite{Extra-sensory perception} to guess what the name of the
closing parenthesis should be. The grammar for XESP is given in
Figure~\ref{fig:xesp} along with the translation to XML documents.
From now on, we use the XESP syntax in our examples.

@@

The web server's job is simple. It runs in a loop, accepting a single
connection, setting up an initial environment $\G$ for evaluation
(which associates the symbols {\sf request.url}, {\sf request.ip} and
{\sf request.time} with the appropriate values), and then evaluating
the document contained at the head of the key ``main'' in the
repository. The result of that evaluation is sent back to the web
browser as a string.\z\footnote{The HTTP result code may be modified
if, for example, the resulting document is a HTTP redirect to another
URL.} The server knows nothing else about how Wikiplia works.

@@

The revision control system stores XESP documents and their history.
This is mostly a straightforward implementation of the imperative {\tt
cvs} judgments given in Section~\ref{sec:calculus}. Unlike typical
revision control systems, we need to support very large numbers of
revisions with small edits (since every change is saved), so the
implementation is engineered to make storing small revisions very
cheap. Particularly, revisions are aggressively compressed by only
storing the newest revision directly, and then a series of difference
``plans'' that describe how to get the next older revision from the
current one. We compute optimal plans using an efficient minimal edit
distance\wcite{Levenshtein distance} calculation at the
token\wcite{Token (parser)} level. As of revision~\currentrevision,
the database is only 236 kilobytes.\z\wcite{Kilobyte}

@@

The key ``main'' of Wikiplia is responsible for decomposing the URL
and acting upon it however is appropriate. The goal of the
bootstrapping process is to make Wikiplia
self-sufficient\wcite{Self-sufficiency} in the sense that the language
can be edited from the web site implemented by ``main.'' To do so, the
web site needs to be able to present the user with an edit box
containing the current source of the ``main'' key and the ability to
save his code into the database, overwriting the ``main'' key, in
order to add functionality.

The initial implementation of ``main'' is provides for the simple
ability to edit, save, and view the current version of keys in the
repository. These three actions are encoded as the URLs {\tt
/edit/key}, {\tt /save/key}, and {\tt /view/\_/key}. The view action
is straightforward. The edit action displays an HTML
textarea\wcite{Text box} containing the current value of the key and a
button that submits the changes to the {\tt save} url. The save action
is the most complex. First, the submitted document is saved as the new
version of the key. Then, if the key is of the form {\tt
base.extension}, the database is checked to see if there is a key
called {\tt extension:compile}. If so, the XESP document that is there
is applied to the input document (a string) to produce a document that
is saved at the key {\tt base}. This allows us to develop languages
that are automatically compiled when saved.

@@

\begin{figure*}[t]
\begin{center}
% \includegraphics[width=0.92\textwidth]{ss_history}
\end{center}
\caption{Screenshot of the history information for {\tt main-go.w}.
} \label{fig:sshistory}
\end{figure*}

@@

The editing interface implemented has been enhanced greatly;
Figure~\ref{fig:sshomepage} shows a screenshot of the main page. The
various views of a key are shown with a series of tabs at the top of
each page. Each user has a home page named after his IP
address,\z\wcite{IP address} which he can use to catalogue his
interests. Various warnings help the user, for example, if he tries to
edit a page that was generated by compiling some source code, a
warning message suggests that he may wish to edit the source code
instead. An ornate logo in SVG\wcite{Scalable Vector Graphics} adorns
the page, and a sidebar provides quick access to the site's features.
The logo and graphics for the site are stored in the repository; the
new {\tt raw} and {\tt typed} actions allows access to these resources
over HTTP so that they may be freely modified.\z\footnote{Because
Wikiplia is dogmatically forward-looking, the graphics require SVG
support in the browser and support for the {\tt data:} URL format, a
combination only found in the newest versions of the Mozilla Firefox.
Wikiplia is compatible with incompatible browsers, however, simply
displaying a crapified version of each page.} At revision 228 support
for metadata was added for each page; a {\tt history} tab now shows
the date, revision number, IP address, and edit summary for each
change to a key (Figure~\ref{fig:sshistory}).

@@

Wikiplia is not just a programming language; it needs facilities for
editing pages that are human readable as well. This can be used to
edit documentation for programming languages, to modify the home page
to tout new developments, to modify the software license
(Section~\ref{sec:license}) or to deface other user's home pages. For
this we provide a Wiki-like\wcite{Wiki} syntax that allows for the
authorship of such pages and easy linking between them. Like other
Wikis, a link to a page that does not exist is colored red, to alert
the user to the opportunity to stake out cyberspace real estate. This
syntax is compiled to XESP documents via the {\tt page} language; the
resulting documents are active in the sense that they check the status
of linked pages to report the correct color on every page load.

@@

Though Wikipedia in its current form is a usable general-purpose
programming language, work remains to be done for it to reach its full
potential. For one, it needs a vibrant community of contentious and
hubristic editors hiding behind anonymous IP addresses boldly
asserting half-baked syntactic extensions or enforcing superficial
style preferences, gritting their teeth while typing and clicking
white knuckled in a kind of road rage\wcite{List of rages} created by
the dehumanizing semantic markup by which they are forced to
communicate.

@@

We also seek to improve the languages. The language {\tt w} needs many
more features to speed development: the parenthesis-based XESP syntax
should eventually be replaced by a pleasant concrete syntax, if we can
get around to it before too much code is written in XESP. A type
system\wcite{Type system} is not planned, because type systems
restrict Freedom~\fexpression, the freedom of expression. However, we
should seek to make Wikiplia as multi-paradigm as possible (again,
freedom from discrimination on the basis of paradigm\wcite{paradigm}
orientation), supporting OOPs-oriented
programming,\z\wcite{Object-oriented programming} aspect-oriented
programming,\z\wcite{Aspect-oriented programming} duck-oriented
typing,\z\wcite{Duck typing} orientation-oriented
orienteering,\z\wcite{Orienteering} Orient-oriented
Programming,\z\wcite{The Orient} etc.

@@

The author doesn't think much of musicals,\z\wcite{Musical theatre} to
be perfectly\wcite{Thomas Aquinas} honest,\z\wcite{Honesty} so those
are basically a no-go. He also feels that the metric
system\wcite{Metric system} and time zones\wcite{Time zone} are pretty
overrated. Ketchup\wcite{Ketchup} on eggs\wcite{Egg (food)} is
gross,\z\wcite{Gross} but not quite as gross as foie gras\wcite{Foie
gras}, which more or less has the word ``gross'' in its name so
duh.\z\wcite{Duh}

>>>>

\vspace{1em}
{\noindent \small {\bf Keywords}:
  Freedom, programming language, software license, wiki, XML, 
  weatherproof footwear and their fastening mechanisms,
  hyper-driven devices
}

@@

\paragraph{Freedom \ftobe: Free to be You and Me.}
The development of a programming language should not be confined to
the bearded academic elite, gazing down upon the programmer fiefs from
their stratospheric ivory towers. Wikiplia is the free programming
language that {\em anyone} can edit: from bearded academic
elite\wcite{Tenure} to congressional staffers\wciteshort{Congressional
staffer edits to Wikipedia} to nameless Slashdot\wcite{Slashdot}
trolls. Wikiplia's WWW-based approach means that an Internet
connection and compatible WWW hypertext browser is all that's needed
to begin on the life-changing journey of programming language design.

@@

The hallmark of Free software is the GNU General Public License. It is
a hereditary license that requires that (a) source code be distributed
with the program and (b) modified versions of the program also be
licensed under the GPL. The intention is that anyone receiving the
software can exercise Freedom~\ftinker\ by understanding the source
code and modifying it to suit his needs. Clearly any source code will
not do: an obfuscated\wcite{Obfuscated code} version of the source
code cannot be easily understood or modified, even though it is
technically ``source code.'' The GPL therefore legally defines source
code as the ``preferred form'' for ``making modifications.''

@@

After this scenario, K and R believe they are working on the same
programming language---after all, it has the same source code---but
their minor bifurcation in development history means that they have
forever different meanings of the \verb+'\c'+ extension. This mistake
is likely to go unnoticed for some time, and until it is resolved, the
meaning of the \verb+'\c'+ extension is firmly enslaved in the
bipartite penitentiary of double entendre, yearning to be
free\,\ldots\z\wcite{Information wants to be free}

@@

Based on these scenarios we conclude that extant social measures such
as revision control conventions are not enough to guarantee freedom.
Even if we think these situations are implausible in the hands of
well-intentioned, well-mannered and capable%
%
\wcite{Year 2000 problem}$^,$%
\wciteshort{Northeast Blackout of 2003}$^,$%
\wcite{Ariane 5 Flight 501}$^,$%
\wcite{Mars Climate Orbiter}$^,$%
\wcite{Mars Polar Lander}$^,$%
\wcite{Mars Rover}$^,$%
\wcite{Mars Pathfinder}$^,$%
\wcite{Gripen\#Crashes}$^,$%
\wcite{missingno.}$^,$%
\wcite{XSS}$^,$%
\wcite{Buffer overflow}$^,$%
\wcite{Therac-25}$^,$%
\wcite{Lothar (storm)}$^,$%
\wcite{Mariner 1}$^,$%
\wcite{Code Red worm}$^,$%
\wcite{SQLSlammer}$^,$%
\wcite{Sandstorm (vehicle)}$^,$%
\wcite{Samy (XSS)}$^,$%
\wcite{Pentium FDIV bug}$^,$%
\wcite{MIM-103 Patriot}$^,$%
\wcite{Windows 95}$^,$%
\wcite{Morris (computer worm)}$^,$%
\wcite{USS Yorktown (CG-48)}
%
software engineers, we wish for our software to remain free even when
in the hands of nefarious and crafty factions who would seek to
fracture\wcite{Fork (software development)} our free software
community. We therefore need a technological and legal solution that
forces the entire development history to be available. Keeping with
Freedom~\fusa, we call this technology and its license {\em Total
Information Awareness} after the successful project of the US
Government with the same name.\z\wcite{Information Awareness Office}

@@

The remainder of this paper proceeds as follows. We first present in
Section~\ref{sec:calculus} the design of the initial Wikiplia system,
which is used to bootstrap the rest of Wikiplia. We then discuss the
current state of Wikiplia as of revision \currentrevision\ in
Section~\ref{sec:current}. We explain the freedom-preserving TIA
Public License in Section~\ref{sec:license}. We conclude with a
discussion of unrelated work and plans for the future~\ref{sec:conc}.

@@

\begin{figure*}[htp]
\[\begin{array}{cc}
% evaluation of lists
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf list}\ \vec{X} \m \tag{list}{\vec{X}}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf cons}\ X\ \tag{list}{\vec{X}} \m \tag{list}{X\ \vec{X}}}{}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf lambda}\ \tag{symbol}{s}\ X \m \tag{closure}{\G\ s\ X}}{}} \\[1em]
\infer{\G \vdash \rate \tag{closure}{\G'\ s\ X}\ \vec{X} \m X'}{\G',s = \tag{list}{\vec{X}} \vdash \eval X \m X'} &
\infer{\G \vdash \rate \tag{prim}{\sf eval}\ X \m X'}{\G \vdash \eval X \m X'} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf xcase}\ \tag{list}{}\ X\ \vec{X} \m X'}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{list}{X_h \vec{X_t}}\  X_0\ \\
                  & \tag{list}{\tag{symbol}{s_h}\ \tag{symbol}{s_t}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_h = X_h, s_t = X_t \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{quote}{X}\  X_0\ X_1\ \\
                  & \tag{list}{\tag{symbol}{s_q}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_q = X \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{string}{s}\  X_0\ X_1\ X_2\ 
                    X\ \vec{X} \m X' \\
\end{array}
}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{int}{i}\  X_0\ X_1\ X_2\ X_3\ 
                    X\ \vec{X} \m X' \\
\end{array}
}{\G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{symbol}{s}\  X_0\ X_1\ X_2\ X_3\ X_4\ \\
                  & \tag{list}{\tag{symbol}{s_b}\ X_b}\ \vec{X} \m X' \\
\end{array}
}{\G, s_b = \tag{string}{s} \vdash \eval X_b \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{
\begin{array}{r@{}l}
  \G \vdash \rate & \tag{prim}{\sf xcase}\ \tag{t}{\vec{X_t}}\  X_0\ X_1\ X_2\ X_3\ X_4\ X_5\ X \vec{X} \m X' \\
\end{array}
}{t = {\tt prim} \,\,or\,\, {\tt closure} & \G \vdash \eval X \m X'}} \\[1em]
\multicolumn{2}{c}{\infer{\G \vdash \rate \tag{prim}{\sf quote}\ {X} \m \tag{list}{X}}{}} \\[1em]
%
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf head}\ \tag{string}{s} \m X
}{{\tt cvs\ checkout}\ s = X}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf read}\ \tag{string}{s}\ \tag{int}{i} \m X
}{{\tt cvs\ checkout\ -\!r}\ i\ s = X}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf insert}\ \tag{string}{s} X \m \tag{int}{i}
}{{\tt cvs\ commit}\ s\ X = i}} \\[1em]
\multicolumn{2}{c}{\infer{
\G \vdash \rate \tag{prim}{\sf history}\ \tag{string}{s} \m \tag{list}{\vec{X}}
}{{\tt cvs\ log}\ s = \vec{X}}} \\[1em]

@@

  \begin{tabular}{r@{\,\,=\,\,}l}
   \lb {\tt(} $E_1\ \ldots\ E_n$ {\tt)}\rb & $\tag{list}{ \lb E_1 \rb\ \ldots\ \lb E_n \rb}$ \\
   \lb {\tt"}$s${\tt"}\rb & $\tag{string}{s}$ \\
   \lb $n$\rb & $\tag{int}{n}$ \\
   \lb $sym$\rb & $\tag{symbol}{sym}$ \\
   \lb {\tt'}$E$\rb & $\tag{quote}{\lb E\rb}$ \\
  \end{tabular}

@@

XML documents are universally parseable.\z\wcite{Parsing} However,
they are difficult to write and read. Therefore, as usual\wcite{RELAX
NG} we create a new syntax by which humans can write and read
documents and which the computer automatically parses and converts to
the easily parseable XML syntax and back to the new syntax, reducing
complexity.\z\footnote{For efficiency, the Wikiplia implementation is
optimized to lazily perform these translations, so that the document
is never represented in XML form.} This compact syntax is based upon
parentheses rather than tags: The XML document $\tag{list}{X_1\ X_2}$
is instead written ${\tt(}X_1\ X_2{\tt)}$. Note that the closing
parenthesis is not {\em named} as in XML, which makes parsing
difficult because the computer must {\em guess} which parenthesis
belongs to which other parenthesis. We therefore call this compact
syntax XESP because it basically {\em reads the programmer's
mind}\,\wcite{Extra-sensory perception} to guess what the name of the
closing parenthesis should be. The grammar for XESP is given in
Figure~\ref{fig:xesp} along with the translation to XML documents.
From now on, we use the XESP syntax in our examples.

@@

The web server's job is simple. It runs in a loop, accepting a single
connection, setting up an initial environment $\G$ for evaluation
(which associates the symbols {\sf request.url}, {\sf request.ip} and
{\sf request.time} with the appropriate values), and then evaluating
the document contained at the head of the key {\tt main} in the
repository. The result of that evaluation is sent back to the web
browser as a string.\z\footnote{The HTTP result code may be modified
if, for example, the resulting document is a HTTP redirect to another
URL.} The server knows nothing else about how Wikiplia works.

@@

The revision control system stores XESP documents and their history.
This is mostly a straightforward implementation of the imperative {\tt
cvs} judgments given in Section~\ref{sec:calculus}. Unlike typical
revision control systems, we need to support very large
numbers\footnote{The implementation also supports revision numbers of
arbitrary size.} of revisions with small edits (since every change is
saved), so the implementation is engineered to make storing small
revisions very cheap. Particularly, revisions are aggressively
compressed by only storing the newest revision directly, and then a
series of difference ``plans'' that describe how to get the next older
revision from the current one. We compute optimal plans using an
efficient minimal edit distance\wcite{Levenshtein distance}
calculation at the token\wcite{Token (parser)} level. As of
revision~\currentrevision, the database is only 236
kilobytes.\z\wcite{Kilobyte}

@@

The key {\tt main} of Wikiplia is responsible for decomposing the URL
and acting upon it however is appropriate. The goal of the
bootstrapping process is to make Wikiplia
self-sufficient\wcite{Self-sufficiency} in the sense that the language
can be edited from the web site implemented by {\tt main}. To do so, the
web site needs to be able to present the user with an edit box
containing the current source of the {\tt main} key and the ability to
save his code into the database, overwriting the {\tt main} key, in
order to add functionality.

The initial implementation of {\tt main} is provides for the simple
ability to edit, save, and view the current version of keys in the
repository. These three actions are encoded as the URLs {\tt
/edit/key}, {\tt /save/key}, and {\tt /view/\_/key}. The view action
is straightforward. The edit action displays an HTML
textarea\wcite{Text box} containing the current value of the key and a
button that submits the changes to the {\tt save} url. The save action
is the most complex. First, the submitted document is saved as the new
version of the key. Then, if the key is of the form {\tt
base.extension}, the database is checked to see if there is a key
called {\tt extension:compile}. If so, the XESP document that is there
is applied to the input document (a string) to produce a document that
is saved at the key {\tt base}. This allows us to develop languages
that are automatically compiled when saved.

@@

\begin{figure*}[t]
\begin{center}
\includegraphics[width=0.92\textwidth]{ss_history}
\end{center}
\caption{Screenshot of the history information for {\tt main-go.w}.
} \label{fig:sshistory}
\end{figure*}

@@

The editing interface implemented has been enhanced greatly;
Figure~\ref{fig:sshomepage} shows a screenshot of the main page. The
various views of a key are shown with a series of tabs at the top of
each page. Each user has a home page named after his IP
address,\z\wcite{IP address} which he can use to catalogue his
interests. Various warnings help the user, for example, if he tries to
edit a page that was generated by compiling some source code, a
warning message suggests that he may wish to edit the source code
instead. An ornate logo in SVG\wcite{Scalable Vector Graphics} adorns
the page, and a sidebar provides quick access to the site's features.
The logo and graphics for the site are stored in the repository; the
new {\tt raw} and {\tt typed} actions allow access to these resources
over HTTP so that they may be freely modified.\z\footnote{Because
Wikiplia is dogmatically forward-looking, the graphics require SVG
support in the browser and support for the {\tt data:} URL format, a
combination only found in the newest versions of the Mozilla Firefox.
Wikiplia is compatible with incompatible browsers, however, simply
displaying a crapified version of each page.} At revision 228 support
for metadata was added for each page; a {\tt history} tab now shows
the date, revision number, IP address, and edit summary for each
change to a key (Figure~\ref{fig:sshistory}).

@@

Wikiplia is not just a programming language; it needs facilities for
editing pages that are human readable as well. This can be used to
edit documentation for programming languages, to modify the home page
to tout new developments, to modify the software license
(Section~\ref{sec:license}) or to deface other user's personal pages.
For this we provide a Wiki-like\wcite{Wiki} syntax that allows for the
authorship of such pages and easy linking between them. Like other
Wikis, a link to a page that does not exist is colored red, to alert
the user to the opportunity to stake out cyberspace real estate. This
syntax is compiled to XESP documents via the {\tt page} language; the
resulting documents are active in the sense that they check the status
of linked pages to report the correct color on every page load.

@@

Though Wikiplia in its current form is a usable general-purpose
programming language, work remains to be done for it to reach its full
potential. For one, it needs a vibrant community of contentious and
hubristic editors hiding behind anonymous IP addresses boldly
asserting half-baked syntactic extensions or enforcing superficial
style preferences, gritting their teeth while typing and clicking
white knuckled in a kind of road rage\wcite{List of rages} created by
the dehumanizing semantic markup by which they are forced to
communicate.

@@

We also seek to improve the languages. The language {\tt w} needs many
more features to speed development: the parenthesis-based XESP syntax
should eventually be replaced by a pleasant concrete syntax, if we can
get around to it before too much code is written in XESP. A type
system\wcite{Type system} is not planned, because type systems
restrict Freedom~\fexpression, the freedom of expression. However, we
should seek to make Wikiplia as multi-paradigm as possible (again,
freedom from discrimination on the basis of paradigm\wcite{paradigm}
orientation), supporting OOPs-oriented
programming,\z\wcite{Object-oriented programming} aspect-oriented
programming,\z\wcite{Aspect-oriented programming} duck-oriented
typing,\z\wcite{Duck typing} orientation-oriented
orienteering,\z\wcite{Orienteering} Orient-oriented
programming,\z\wcite{The Orient} etc.

@@

The author doesn't think\wcite{Thought} much of
musicals,\z\wcite{Musical theatre} to be perfectly\wcite{Thomas
Aquinas} honest,\z\wcite{Honesty} so those are basically\wcite{BASIC}
a no-go. He also feels that the metric system\wcite{Metric system} but
paradoxically\wcite{Pardadox} also time zones\wcite{Time zone} are
pretty overrated. Ketchup\wcite{Ketchup} on eggs\wcite{Egg (food)} is
gross,\z\wcite{Gross} but not quite as gross as foie gras\wcite{Foie
gras}, which more or less has the word\wcite{Word} ``gross'' in its
name\wcite{Name} so duh.\z\wcite{Duh}
